---
title: "Canada bikes like the Dutch"
output:
  word_document: default
editor_options:
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


#Part A: How much do Canadians cycle per day

We will determine how much Canadians cycle per day by calculating the average daily energy expenditure from cycling as a form of transportation.

We will use 3 variables in our data set which are based off of the 2013/2014 CCHS PUMF and measure cycling as a form of active transportation:

- PAC_8: In the past 3 months did you cycle to and from work or school?
- PAC_8A: How many times did you cycle to and from work or school, in th past 3 months?
- PAC_8B:	How much time did you spend on each occasion?


**Step 1:** Import your data set to R.

```{r}
paData <- read.csv("CCHS-2013-1-percent.csv", header = TRUE)
```

**Reminder:** Although the CCHS-2013-1-percent.csv is based on the 2013/2014 Canadian Community Health Survey Public Use Microdata File, CCHS-2013-1-percent.csv is a completely synthetic data set and can only be used for examplary purposes.

Assumption #1: we assume individuals 65 or older are retired and therefore not cycling to/from work/school. Therefore, we will only evaluate individuals that are < 65 years old.

**Step 2:** Filter the data: < 65. The CCHS data set has categories for age, therefore we will filter out all age categories that are >=65 years old.
```{r}
library(dplyr)

paData <- paData %>%
  filter(DHHGAGE <= 12)
```

We will first find the average of cycling per day for individuals that **did** cycle to/from work/school (PAC_8, answer 1) only:


**Step 3:** Find the average number of times, that an individual cycled to/from work/school in the past 3 months
```{r}
#a. Exclude the missing data - all individuals that did not cycle in the past 3 months: do not cycle to/from work school, do not know, refused to answer, or did not answer
#Make 996, 997, 998, and 999 to N/A for PAC_8A
paData$PAC_8A[paData$PAC_8A == 996] <- NA
paData$PAC_8A[paData$PAC_8A == 997] <- NA
paData$PAC_8A[paData$PAC_8A == 998] <- NA
paData$PAC_8A[paData$PAC_8A == 999] <- NA

#b. Calculate the average number of times that an individual cycled to/from work/school (mean of PAC_8A)
(freqCycle <- mean(paData$PAC_8A, na.rm = TRUE))
```


**Step 4:** Find the average time spent cycling to/from work/school, in the past 3 months
```{r}
#a.	Exclude the missing data - all individuals that did not cycle in the past 3 months: do not cycle to/from work school, do not know, refused to answer, or did not answer

#Make 6, 7, 8, and 9 to N/A  for PAC_8B:
paData$PAC_8B[paData$PAC_8B == 6] <- NA
paData$PAC_8B[paData$PAC_8B == 7] <- NA
paData$PAC_8B[paData$PAC_8B == 8] <- NA
paData$PAC_8B[paData$PAC_8B == 9] <- NA

#b.	Calculate the mean time spent cycling (mean of PAC_8B) in minutes
(timeCycle <- 15 * mean(paData$PAC_8B, na.rm = TRUE))

#d. convert from minutes to hours
(timeHourCycle <- timeCycle/60)
```


**Step 5:**	Calculate the average duration of cycling per day
```{r}
(durationCycle <- (timeHourCycle*freqCycle)/(30.42*3))
```


**Step 6:** Calculate the average daily energy expenditure (MET-hours) for cycling to/from work/school per day
```{r}
(metHoursCycle <- (durationCycle*5.8))
```
We now have our average daily energy expenditure (MET-hours) for Canadians that **DO** cycle to/from work/school per day.


Now we will account for the entire population, including those that did not cycle to/from work school.

**Step 7:**	Determine the average daily expenditure for all individuals in the population not only those that cycled to/from work/school per day
```{r}
#a. porpotion of inividuals that do cycle to/from work/school
(paTableCycle <- table(paData$PAC_8))
(indivCycle <- paTableCycle[1]/(sum(paTableCycle)))

#b.	Calculate the average daily energy expenditure from cycling to/from work/school (MET-hours) for the entire population
(metHoursCycleAll <- metHoursCycle*indivCycle)
```

On average a Canadian male only gets 0.016 MET-hours/day from cycling to/from school/work.



#Part 2: How much do the Dutch cycle/day for transportation purposes

Using data collected as part of the Dutch National Travel Survey (2010 – 2012), Fisherman, 2015 (doi:10.1371/journal.pone.0121871 - Table S1) reported the average additional daily energy expenditure from cycling as a form of transportation for the Dutch: males: 1.3 MET-hours, and females: 1.4 MET-hours.

The average Dutch daily energy expenditure from cycling as a form of transportation for both sexes is then 1.35 MET-hours.



#Part 3: Calculate the difference between Dutch and Canadian daily cycling levels

Canadians need to increase their daily energy expenditure from cycling from 0.07 MET-hours to 1.35 MET-hours, in order to be like the Dutch.
```{r}
absChange <- 1.35 - metHoursCycleAll
```



#Part 4: Predict how Canadian life expectancy would change if Canadians biked like the Dutch.

Use the Project Big Life Planning Tool.

