---
title: "Healty Cities""
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#Part A: Identify the healthiest region

Using the Project Big Life Planning Tool calculate the Summary life expectancy for each region in Canada, and download the summary data set. 

**Reminder:** Although the data.sample.csv is based on the 2013/2014 Canadian Community Health Survey Public Use Microdata File, data.sample.csv is a completely synthetic data set and can only be used for examplary purposes.

1. Identify the healthiest region. 
```{r}
#Import the population summary data set
  #data <- read.csv("pop.data.sample.csv")

#Identify the healthiest region
  #pop.data.sample[pop.data.sample$SummaryLE == max(pop.data.sample$SummaryLE),]
```

###Part B: Determine the health behaviours of the healthiest region and the Ottawa region. 

Calculate the average of each health behaviour in the healthiest region: XXX and the Ottawa region.

1. Import your original data set to R. 
```{r}
 data <- read.csv("data.sample.csv", header = TRUE)
```

2. Subset your data for the healthiest region XXX and the Ottawa region. 
```{r}
    library(dplyr)

    Ottawa <- data %>%
      filter(GEODPMF == 35951)
    
    XXX <- data %>%
      filter(GEODPMF == 59922)
```

3. Calculate the average/prevalence of each health behaviour for the XXX and Ottawa region.

**Smoking**

Calculate the prevalance of current smokers in the XXX and Ottawa region.
```{r}
  smokersOttawa <- table(Ottawa$SMKDSTY)
  prevSmokersOttawa <- round(prop.table(smokersOttawa),2)
  prevSmokersOttawa[1]

  smokersXXX <- table(XXX$SMKDSTY)
  prevSmokersXXX <- round(prop.table(smokersXXX),2)
  prevSmokersXXX[1]
```


**Physical Activity**

Calculate the average of physical activity (varibale PACDEE) in the XXX and Ottawa region.
a. Exclude the missing data: 'not stated'
```{r}
  Ottawa$PACDEE[Ottawa$PACDEE == 99.9] <- NA
  XXX$PACDEE[XXX$PACDEE == 99.9] <- NA
```

b. Calculate the average daily energy expenditure for individuals in each region who do have a value for PACDEE
```{r}
  paOttawa <- mean(Ottawa$PACDEE, na.rm = TRUE)
  paXXX <- mean(XXX$PACDEE, na.rm = TRUE)
```  


**Alcohol**

Calculate the average of weekly alcohol consumption (varibale ALWDWKY) in the XXX and Ottawa region.

a. Exclude the missing data: 'not stated'
```{r}
  Ottawa$ALWDWKY[Ottawa$ALWDWKY == 999] <- NA
  XXX$ALWDWKY[XXX$ALWDWKY == 999] <- NA
```

b. Change observations: not applicable to 0. 
These are individuals who did not drink alcohol in the past year (represented by observations coded as 996 - NA).
```{r}
  Ottawa$ALWDWKY[Ottawa$ALWDWKY == 996] <- 0
  XXX$ALWDWKY[XXX$ALWDWKY == 996] <- 0
```

c. Calculate the average alcohol consumption for individuals in each region
```{r}
  alcoholOttawa <- mean(Ottawa$ALWDWKY, na.rm = TRUE)
  alcoholXXX <- mean(XXX$ALWDWKY, na.rm = TRUE)
```  


**Diet**

An individual's diet is composed of multiple components including the number of daily servings of: carrot, fruit, juice, potato, salad, and vegetable. 

The more daily servings an individual has of:

- carrots, fruits, salad, and vegetables =  the more healthy the individual, 
- juice = the less healthy the individual,
- potato < 1 = the more healthy the individual,  and
- potato > 1 = the less healthy the individual. 

To compare the average diet of each region


Calculate the average of each diet component for the XXX and Ottawa region.

a. Eliminate observations: not stated

```{r}
  dietOttawa <- Ottawa %>%
    select(FVCDCAR, FVCDFRU, FVCDJUI, FVCDPOT, FVCDSAL, FVCDVEG)
  dietOttawa[dietOttawa == 999.9] <- NA

  dietXXX <- XXX %>%
    select(FVCDCAR, FVCDFRU, FVCDJUI, FVCDPOT, FVCDSAL, FVCDVEG)
  dietXXX[dietXXX == 999.9] <- NA
```

b. Calculate the average of each diet component for XXX and the Ottawa region.
```{r}
  summary(dietOttawa)
  summary(dietXXX)
```

**Comparing health behaviours**

4. Compare the averages of all health behaviours in the Ottawa region to the XXX. When the health behaviour for Ottawa is unhealthier then XXX, record the avg/prevalence of that health behaviour for XXX.


###Part C: What if Ottawa acted like XXX?

Use the Project Big Life Planning Tool to run an intervention scenario where the Ottawa region acts like XXX.